---
title: "Ames Housing ML"
author: "Henry Greeley"
date: "October 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load data
```{r}
library(dplyr)   # gives mutate_if
library(forcats) # gives fct_explicit_na
library(dummies)  #encoding

train <- read.csv("train.csv")

#Replace NA with value 0 or (missing)
train[is.na(train)] <- 0

#replace nas for factors with characters
train = train %>% 
	mutate_if(is.factor, fct_explicit_na)

train = train %>% 
	mutate_if(is.factor, fct_explicit_na, na_level = 'NA')

```

#load prediction data
```{r}

test <- read.csv("test.csv")

#Replace NA with value 0 or (missing)
test[is.na(test)] <- 0

test = test %>% 
	mutate_if(is.factor, fct_explicit_na)

test = test %>% 
	mutate_if(is.factor, fct_explicit_na, na_level = 'NA')

```


#harmonize columns between data frames with dummy encoding for all records

```{r}

traintest = rbind(train[,-ncol(train)], test)

```

#dummy encoding for all variables contained in both training and test sets
```{r}
ttdummy <- dummy.data.frame(traintest)

#fix column names to syntactically valid names
names(ttdummy) <- make.names(names(ttdummy))
#names(ttdummy)
```

#merge the dummy data back to train data
```{r}
train2 <- merge(ttdummy,train[,c(1,ncol(train))], by = "Id")
```


#fit random forest regressor
```{r}
library(randomForest)

ames.rf = randomForest(SalePrice ~ . , data = train2 )

ames.rf
```

```{r}
plot(ames.rf)

```

#connect test data to dummy variables created earlier
```{r}
test2 <- ttdummy[1461:nrow(ttdummy),]

nrow(test2)
```


#Prediction on test set
```{r}
pred <- predict(ames.rf, test2)

#pred
```

#output file for kaggle
```{r}

pred <- data.frame(pred)
modout <- cbind(test2[,1:1],pred)
names(modout) <-  c("Id", "SalePrice")
```

```{r}
write.csv(modout, file = "ameskaggle.csv")
```



#Build shiny app to visually analyze variables efficiently
```{r}
library(shiny)
```

```{r}
# Define UI for application that draws a histogram
ui <- fluidPage(

# Give the page a title
titlePanel("Ames Housing Feature Visualization"),

# Generate a row with a sidebar
sidebarLayout(

# Define the sidebar with one input
sidebarPanel(
selectInput("region", "Feature:",
choices=colnames(tdata)),
hr(),
helpText("Select Items to Analyze")
),

# Create a spot for the barplot
mainPanel(
plotOutput("phonePlot")
)
)
)
```

```{r}
# Define server logic required to draw a histogram
server <- function(input, output) {
  
  # Fill in the spot we created for a plot
  output$phonePlot <- renderPlot({
    
    # Render a barplot
    barplot(table(tdata[,input$region]), 
            main=input$region,
            ylab="Frequency",
            xlab="Feature")
  })
}

# Run the application 
shinyApp(ui = ui, server = server)
```
#Create training and test datasets
```{r}
library(ISLR)
smp_siz = floor(0.75*nrow(tdata)) 
smp_siz
```

```{r}
library(dplyr)   # gives mutate_if
library(forcats) # gives fct_explicit_na

set.seed(13)   # set seed to ensure you always have same random numbers generated
train_ind = sample(seq_len(nrow(tdata)),size = smp_siz)  # randomly select 75% of rows

train = tdata[train_ind,] #training dataset
train[is.na(train)] <- 0

#Change NA to factor or 0

train = train %>% 
	mutate_if(is.factor, fct_explicit_na)

train = train %>% 
	mutate_if(is.factor, fct_explicit_na, na_level = 'NA')


test = tdata[-train_ind,]  #test dataset
test[is.na(test)] <- 0

test = test %>% 
	mutate_if(is.factor, fct_explicit_na)

test = test %>% 
	mutate_if(is.factor, fct_explicit_na, na_level = 'NA')

#test[is.na(test)] <- "NA"

```






